On Secure Delegation of SampBQP with a Classical Client

### 2020-02-09 bqp.tex
Restructure BQP protocols and rewrite proof

\begin{protocol}{1-message $\QPIP_1$ protocol $\PiMF = (\PMF, \VMF)$ for $\BQP$ language $L$}
	\label{proto:BQP}
	Common input: $x\in\set{0,1}^n$. 
	\begin{enumerate}
		\item The verifier generates $\Lambda \leftarrow \cVMF{1}(x)$, where $\Lambda=\Lambda_1\otimes\ldots\otimes\Lambda_m$; each $\Lambda_i$ is a single-qubit $X$ or $Z$ measurement and $m=\poly(n)$ is the length of the prover's certificate state. \Ethan{Switch steps 1 and 2}
		\item The prover generates a certificate state $\rho\leftarrow\cPMF{2}(x)$  and sends $\rho$ to the verifier qubit-by-qubit.
		\item The verifier applies the measurement $\Lambda$ to $\rho$ qubit-by-qubit.
			Let $w \in \set{0,1}^m$ denote the measurement outcome.
			It verifies the measurement outcome and produces a decision $o\leftarrow\cVMF{3}(x, \Lambda, w)$ where $o\in\set{acc, rej}$ and $\cVMF{3}$ is a classical polynomial-time algorithm.
	\end{enumerate}
\end{protocol}

We compile this protocol using Mahadev's $\mathsf{QHE}$ from \cite{mahadev_qfhe} to obtain a blind 1-message $\QPIP_1$ protocol for $\BQP$, again with only $X$ and $Z$ measurements.
The high level idea is to run \cref{proto:BQP} under $\mathsf{QHE}$.
Namely, the verifier encrypts its original input $x$ and sends $y=\mathsf{QHE.Enc}(x)$ to the prover.
The prover then homomorphically evaluate $\cF$ on $y$ to produce $\sigma=\mathsf{QHE.Enc}(\rho)$ \Ethan{Is this actually ok? $\sigma$ is a ciphertext corresponding to $\rho$, but it wasn't encrypted. In fact it's deteriorated due to homomorphic eval} and send it back to the verifier.

@@

\begin{thm}
	\label{decodeMeasureOrder}
	Let $sk$ be some secret key of $\mathsf{QHE}$.
	Let $n\in\bbN$.
	Let $m\in\bbN$ be the length of the ciphertexts under $\mathsf{QHE}$ corresponding to plaintexts of length $n$.
	Let $\Lambda=\Lambda_1\otimes\ldots\otimes\Lambda_n$, where each $\Lambda_i$ is a single-qubit $X$ or $Z$ measurement.

	There exists $\Lambda'=\Lambda'_1\otimes\ldots\otimes\Lambda'_m$ where each $\Lambda_i'$ is a single-qubit $X$ or $Z$ measurement, and polynomial-time algorithm $\widehat{\mathsf{QHE.Dec}}$ such that:
	$$\Lambda\circ\mathsf{QHE.Dec}=\widehat{\mathsf{QHE.Dec}}\circ\Lambda'$$
\end{thm}
\begin{prf}
	This is a proof by construction.
	We require $\Lambda'$ and $\widehat{\mathsf{QHE.Dec}}$ such that $\forall\sigma=(X^xZ^z\rho Z^zX^x, c_{x,z})$,
	$$\Lambda\circ\mathsf{QHE.Dec}(\sigma)=\widehat{\mathsf{QHE.Dec}}\circ\Lambda'(\sigma)$$

	Clearly,
	$$\Lambda\circ\mathsf{QHE.Dec}(\sigma)=\Lambda(\rho)$$

@@

\Ethan{Give the theorem above a corresponding algorithm}

\Cref{proto:BQP} has negligible completeness and soundness.
Now we compile it using $\mathsf{QHE}$ in order to also achieve blindness:

@@

	Let $d=\poly(n)$ be the depth of the circuit corresponding to $\cPMF{2}$.
	\begin{enumerate}
		\item The verifier generates $\Lambda\leftarrow\cVMF{1}(x)$.
			On input $\Lambda$, it computes $\Lambda'$ using \cref{decodeMeasureOrder}.
			Then it generates $(pk, sk, evk)\leftarrow\mathsf{QHE.Keygen}(1^n, 1^d)$ and $y\leftarrow\mathsf{QHE.Enc}_{pk}(x)$.
			It sends $y$ and $evk$ to the prover.
			\Ethan{Do two steps. Keygens first, lambda next. For keygen step, say more in words and describe $d$. Remember to specify $x$ when talking about $d$ and $\cPMF{2}$}
		\item The prover homomorphically evaluates $\cPMF{2}$ on input $y$, generating $\sigma\leftarrow\mathsf{QHE.Eval}(\cPMF{2}, y)$.
			It sends $\sigma$ to the verifier qubit-by-qubit.
		\item The verifier generates $ct\leftarrow\Lambda'(\sigma)$,
			$w\leftarrow\widehat{\mathsf{QHE.Dec}_{sk}}(ct)$,
			and then $o\leftarrow\cVMF{3}(x, \Lambda, w)$.
			It outputs $o$.
	\end{enumerate}
\end{protocol}

\Ethan{Change steps orders above. 1. verifier enc and send y 2. prover sends something back 3. verifier sample lambda}

@@

\begin{thm}
	$\Piblind$ has negligible completeness and soundness, and is blind.
\end{thm}
\begin{prf}
	The completeness follows by inspection.
	For blindness, the only thing the prover sees is $c$ and $evk$, so this follows directly from $\mathsf{QHE}$'s security.

	Suppose for the sake of contradiction that there exists some cheating prover $\Pblindstar$ violating the soundness guarantees of $\Piblind$ 
	\Ethan{Define some $\epsilon$ here. Also ``non-negligible" = ``noticeable"},
	then we can use it to construct a prover $\PMFstar$ violating the soundness guarantees of $\PiMF$.

	\Ethan{This construction is too small for a separate box but maybe still be too big to just write it like this?}

	\Ethan{Add the ``simulate" thing back}

	We define $\PMFstar$ to first generate
	$(pk, sk, evk)\leftarrow\mathsf{QHE.Keygen}$
	and
	$y\leftarrow\mathsf{QHE.Enc}_{pk}(x)$
	It then sends $y$ and $evk$ to $\Pblindstar$ and gets $\sigma$ back.
	Finally, it decrypts
	$\rho\leftarrow\mathsf{QHE.Dec}_{sk}(\sigma)$
	and sends it to $\VMF$ qubit-by-qubit.

    \Ethan{Rewrite the following proof again; write in terms of interactions this time. See google docs. Same success probability. Compare interactions}

	We now prove that $\PMFstar$ in fact violates the soundness of $\PiMF$.
    By assumption, $\Vblind$ has a non-negligible probability to accept after receiving $\sigma$.
	In other words, let $\Lambda\leftarrow\cVMF{1}$, and define $\Lambda'$ by \cref{decodeMeasureOrder},
    \Ethan{Give theorem 3.2 an algorithm}
	then with non-negligible probability
	$$\cVMF{3}(x, \Lambda, \widehat{\mathsf{QHE.Dec}_{sk}}(\Lambda'(\sigma))=acc$$
	which is equivalent to
	$$\cVMF{3}(x, \Lambda, \Lambda(\rho))=acc$$
	That is, $\PMFstar$ violates the soundness of $\PiMF$.
\end{prf}

@@

Finally, we prove Thm 3.1 by applying Mahadev's measurement protocol~\cite{mahadev_delegation} to Protocol 3 to compile it into a $\QPIP_0$ protocol $(P', V')$ \Ethan{fix name}.
Note that Mahadev's measurement protocol can compile any $\QPIP_1$ protocol with X, Z measurement and negligible soundness to a $\QPIP_0$ protocol, and Protocol 3 is a $\QPIP_1$ protocol with X, Z measurement and negligible soundness.
Thus the resulting protocol $(P', V')$ is a O(1)-round $\QPIP_0$ protocol with soundness error $\frac{3}{4} + \negl(n)$

@@

\begin{protocol}{\Cref{proto:BlindBQP} compiled using Mahadev's measurement protocol \Ethan{Give this a name}}
	\label{proto:QPIP0BQP}
	\begin{enumerate}
		\item The verifier generates $\Lambda, \Lambda', pk, sk, evk, y$ the same way as in \cref{proto:BlindBQP}. \Ethan{Need more notation? But how?}
			Then, using $\Lambda'$, it generates $k, td$.
			It sends $k, evk, y$ to the prover.
		\item The prover, on inputs $k, evk, y$, generates classical ``commitment" $m$ and internal state $\ket{st}$.
			It sends $m$ to the verifier.
		\item The verifier sends the prover a public coin $c\in\set{0,1}$.
		\item The prover, on inputs $c$ and $\ket{st}$, generates classical $a$ and sends it to the verifier.
		\item The verifier, on inputs $k, td, m, c, a, sk$, generates and outputs either $acc$ or $rej$.
	\end{enumerate}
\end{protocol}

>>>>

\begin{protocol}{1-message $\QPIP_1$ protocol $\PiMF = (\PMF, \VMF)$ for $\BQP$ language $L$}
	\label{proto:BQP}
	Common input: $x\in\set{0,1}^n$. 
	\begin{enumerate}
		\item The prover generates a certificate state $\rho\leftarrow\cPMF{1}(x)$  and sends $\rho$ to the verifier qubit-by-qubit.
		\item The verifier generates $\Lambda \leftarrow \cVMF{2}(x)$, where $\Lambda=\Lambda_1\otimes\ldots\otimes\Lambda_m$;
			each $\Lambda_i$ is a single-qubit $X$ or $Z$ measurement and $m=\poly(n)$ is the length of the prover's certificate state.
			It applies $\Lambda$ to $\rho$ qubit-by-qubit.
			Let $w \in \set{0,1}^m$ denote the measurement outcome.
			It verifies the measurement outcome and produces a decision $o\leftarrow\cVMF{3}(x, \Lambda, w)$ where $o\in\set{acc, rej}$ and $\cVMF{3}$ is a classical polynomial-time algorithm.
	\end{enumerate}
\end{protocol}

$\PiMF$ has negligible completeness and soundness.

We compile this protocol using Mahadev's $\QHE$ from \cite{mahadev_qfhe} to obtain a blind 1-message $\QPIP_1$ protocol for $\BQP$, again with only $X$ and $Z$ measurements.
The high level idea is to run $\PiMF$ under $\mathsf{QHE}$.
Namely, the verifier encrypts its original input $x$ and sends $y=\mathsf{QHE.Enc}(x)$ to the prover.
The prover then homomorphically evaluate $\cF$ on $y$ to produce $\sigma$ which is a ciphertext corresponding to $\rho$, and sends $\sigma$ to the verifier.

@@

\Ethan{TODO PROOFREAD ME}
\begin{thm}
	\label{decodeMeasureOrder}
	Let $n\in\bbN$.
	Let $m\in\bbN$ be the length of the ciphertexts under $\mathsf{QHE}$ corresponding to plaintexts of length $n$.

	There exists efficient classical poly-time algorithms $\cM_\QHE$ and $\widehat{\mathsf{QHE.Dec}}$ s.t.
	for all $\Lambda=\Lambda_1\otimes\ldots\otimes\Lambda_n$ where each $\Lambda_i$ is a single-qubit $X$ or $Z$ measurement:
	\Ethan{How to specify the existence of $sk$?}
	$$\cM_\QHE(\Lambda)=\Lambda'=\Lambda'_1\otimes\ldots\otimes\Lambda'_m$$
	where each $\Lambda_i'$ is a single-qubit $X$ or $Z$ measurement, and
	$$\Lambda\circ\mathsf{QHE.Dec}=\widehat{\mathsf{QHE.Dec}}\circ\Lambda'$$
\end{thm}
\begin{prf}
	This is a proof by construction.
	We require $\Lambda'$ and $\widehat{\mathsf{QHE.Dec}}$ such that $\forall\sigma=(X^xZ^z\rho Z^zX^x, c_{x,z})$,
	$$\Lambda\circ\mathsf{QHE.Dec}(\sigma)=\widehat{\mathsf{QHE.Dec}}\circ\Lambda'(\sigma)$$
	Clearly,
	$$\Lambda\circ\mathsf{QHE.Dec}(\sigma)=\Lambda(\rho)$$

@@

Now we state our construction to compile $\PiMF$ using $\mathsf{QHE}$:

@@

	Let $d=\poly(n)$ be the depth of the circuit corresponding to $\cPMF{1}$.
	\begin{enumerate}
		\item The verifier generates $(pk, sk, evk)\leftarrow\mathsf{QHE.Keygen}(1^n, 1^d)$, where $d$ is the depth of the circuit corresponding to $\cPMF{1}(x)$.
			It then generates and sends $ct_x\leftarrow\mathsf{QHE.Enc}_{pk}(x)$ to the prover.
			It also sends $evk$ to the prover.
		\item The prover homomorphically evaluates $\cPMF{1}$ on input $ct_x$, generating $\sigma\leftarrow\mathsf{QHE.Eval}_{evk}(\cPMF{1}, ct_x)$.
			It sends $\sigma$ to the verifier qubit-by-qubit.
		\item The verifier generates $\Lambda\leftarrow\cVMF{2}(x)$,
			$\Lambda'\leftarrow\cM_\QHE$,
			$ct_w\leftarrow\Lambda'(\sigma)$,
			$w\leftarrow\widehat{\mathsf{QHE.Dec}_{sk}}(ct_w)$,
			and $o\leftarrow\cVMF{3}(x, \Lambda, w)$.
			It outputs $o$.
	\end{enumerate}
\end{protocol}

@@

\begin{thm}
	$\Piblind$ has negligible completeness and soundness, and is blind.
\end{thm}
\begin{prf}
	The completeness follows by inspection.
	For blindness, the only thing the prover sees is $c$ and $evk$, so it follows directly from $\mathsf{QHE}$'s security.

	Suppose for the sake of contradiction that there exists some cheating prover $\Pblindstar$ with some noticeable acceptance probability $\varepsilon$.

	Then we can construct a prover $\PMFstar$ violating the soundness guarantees of $\PiMF$ by simulating $(\Vblind(x), \Pblindstar)$ as below:

	We define $\PMFstar$ to first generate
	$(pk, sk, evk)\leftarrow\mathsf{QHE.Keygen}(1^n, 1^d)$
	and
	$y\leftarrow\mathsf{QHE.Enc}_{pk}(x)$
	It then sends $y$ and $evk$ to $\Pblindstar$ and gets $\sigma$ back.
	Finally, it decrypts
	$\rho\leftarrow\mathsf{QHE.Dec}_{sk}(\sigma)$
	and sends it to $\VMF$ qubit-by-qubit.

	We now prove that $\PMFstar$ in fact violates the soundness of $\PiMF$.
	Let $\cP_\blind^*$ denote the action $\Pblindstar$ takes to produce $\sigma$ from $ct_x$.
	Consider the experiment corresponding to $\Piblind$:
	$$(pk, sk, evk)\leftarrow\QGen(1^n, 1^d)$$
	$$ct_x\leftarrow\QEnc_{pk}(x)$$
	$$\sigma\leftarrow\cP_\blind^*(ct_x, evk)$$
	$$\Lambda\leftarrow\cVMF{2}(x)$$
	$$\Lambda'\leftarrow\cM_\QHE(\Lambda)$$
	$$ct_w\leftarrow\Lambda'(\sigma)$$
	$$w\leftarrow\widehat{\QDec_{sk}}(ct_w)$$
	$$o\leftarrow\cVMF(x, \Lambda, w)$$

	Compare it with the experiment corresponding to $\PMFstar$:
	$$(pk, sk, evk)\leftarrow\QGen(1^n, 1^d)$$
	$$ct_x\leftarrow\QEnc_{pk}(x)$$
	$$\sigma\leftarrow\cP_\blind^*(ct_x, evk)$$
	$$\Lambda\leftarrow\cVMF{2}(x)$$
	$$\rho\leftarrow\QDec_{sk}(\sigma)$$
	$$w\leftarrow\Lambda(\rho)$$
	$$o\leftarrow\cVMF(x, \Lambda, w)$$

	By inspection, the only difference between the experiments is how $w$ is obtained from $\sigma$.
	By \cref{decodeMeasureOrder}, the two ways are equivalent.
\end{prf}

@@

Finally, we prove Thm 3.1 by applying Mahadev's measurement protocol~\cite{mahadev_delegation} to Protocol 3 to compile it into a $\QPIP_0$ protocol $(P', V')$ \Ethan{fix name}.
Note that Mahadev's measurement protocol can compile any $\QPIP_1$ protocol with X, Z measurement and negligible soundness to a $\QPIP_0$ protocol, and Protocol 3 is a $\QPIP_1$ protocol with X, Z measurement and negligible soundness.
Thus the resulting protocol $(P', V')$ \Ethan{name?} is a O(1)-round $\QPIP_0$ protocol with soundness error $\frac{3}{4} + \negl(n)$

@@

\begin{protocol}{\Cref{proto:BlindBQP} compiled using Mahadev's measurement protocol \Ethan{Give this a name}}
	\label{proto:QPIP0BQP}
	\begin{enumerate}
		\item The verifier generates $\Lambda, \Lambda', pk, sk, evk, ct_x$ the same way as in \cref{proto:BlindBQP}. \Ethan{Need more notation? But how?}
			Then, using $\Lambda'$, it generates $k, td$.
			It sends $k, evk, ct_x$ to the prover.
		\item The prover, on inputs $k, evk, ct_x$, generates classical ``commitment" $y$ and internal state $\ket{st}$.
			It sends $y$ to the verifier.
		\item The verifier sends the prover a public coin $c\in\set{0,1}$.
		\item The prover, on inputs $c$ and $\ket{st}$, generates classical $a$ and sends it to the verifier.
		\item The verifier, on inputs $k, td, y, c, a, sk$, generates and outputs either $acc$ or $rej$.
	\end{enumerate}
\end{protocol}
